---
title: "The gray ROC"
author: "Giovanni Parmigiani"
date: "2/5/2019"
output:
  pdf_document: default
  html_document: default
---

Script for producing the figures in "Receiver Operating Curves with Indeterminacy". 

Loading data package.

```{r, include=FALSE}
library(curatedOvarianData)
data(GSE32063_eset)
```

Loading gray ROC package

```{r, include=FALSE}
library(grayROC)
```

\newpage

Select gene of interest ("ZNF544").
Make default dotplot (not used in paper)

```{r}
GeneName = "ZNF544"
XX = exp ( exprs(GSE32063_eset)[GeneName,] )
YY = 1 * ( pData(GSE32063_eset)[,"debulking"] == "optimal" )
pdf(file="../PRLettersV2Latex/figures/dotplot.pdf",width=7,height=3)
  plot(XX,YY,xlab=c("ZNF544 Expression"),ylab="CLASS LABEL",ylim=c(-.1,1.1),cex=2)
dev.off()
```

Make didactic boxplot fo Figure 1.

```{r}
source("grayROCdidacticDotplot.R")
pdf(file="../PRLettersV2Latex/figures/didacticDotplot.pdf",width=7,height=7)
  grayROCdidacticDotplot(XX,YY,cL=2.8,cU=3.5,xlab=c("ZNF544 Expression"),ylab="CLASS LABEL")
dev.off()
```
\newpage

Produce gray ROC plots for Figures 2 and 3. The function assumes that low levels of the biomarker predict class 0.

```{r}
ROC = grayROC(XX,YY,0)
fROC = grayROC(XX,YY,.1)
pdf(file="../PRLettersV2Latex/figures/grayROC10set.pdf")
  grayROCplot(fROC,mainTitle="Gamma = .10",set=TRUE,segm=TRUE)
  lines(1-ROC$F0star,1-ROC$F1star)
dev.off()
pdf(file="../PRLettersV2Latex/figures/grayROC10seg.pdf")
  grayROCplot(fROC,mainTitle="Gamma = .10",set=FALSE,segm=TRUE)
dev.off()

ROC = grayROC(XX,YY,0)
fROC = grayROC(XX,YY,.15)
pdf(file="../PRLettersV2Latex/figures/grayROC15set.pdf")
  grayROCplot(fROC,mainTitle="Gamma = .15",set=TRUE,segm=TRUE)
  lines(1-ROC$F0star,1-ROC$F1star)
dev.off()
pdf(file="../PRLettersV2Latex/figures/grayROC15seg.pdf")
  grayROCplot(fROC,mainTitle="Gamma = .15",set=FALSE,segm=TRUE)
dev.off()

ROC = grayROC(XX,YY,0)
fROC = grayROC(XX,YY,.20)
pdf(file="../PRLettersV2Latex/figures/grayROC20set.pdf")
  grayROCplot(fROC,mainTitle="Gamma = .20",set=TRUE,segm=TRUE)
  lines(1-ROC$F0star,1-ROC$F1star)
dev.off()
pdf(file="../PRLettersV2Latex/figures/grayROC20seg.pdf")
  grayROCplot(fROC,mainTitle="Gamma = .20",set=FALSE,segm=TRUE)
dev.off()

ROC = grayROC(XX,YY,0)
fROC = grayROC(XX,YY,.25)
pdf(file="../PRLettersV2Latex/figures/grayROC25set.pdf")
  grayROCplot(fROC,mainTitle="Gamma = .25",set=TRUE,segm=TRUE)
  lines(1-ROC$F0star,1-ROC$F1star)
dev.off()
pdf(file="../PRLettersV2Latex/figures/grayROC25seg.pdf")
  grayROCplot(fROC,mainTitle="Gamma = .25",set=FALSE,segm=TRUE)
dev.off()

ROC = grayROC(XX,YY,0)
fROC = grayROC(XX,YY,.30)
pdf(file="../PRLettersV2Latex/figures/grayROC30set.pdf")
  grayROCplot(fROC,mainTitle="Gamma = .30",set=TRUE,segm=TRUE)
  lines(1-ROC$F0star,1-ROC$F1star)
dev.off()
pdf(file="../PRLettersV2Latex/figures/grayROC30seg.pdf")
  grayROCplot(fROC,mainTitle="Gamma = .30",set=FALSE,segm=TRUE)
dev.off()

```

Make the two files used in Figure 4.

```{r}
fROC = grayROC(XX,YY,.2)
pdf(file="../PRLettersV2Latex/figures/grayZoneSize.pdf")
par(mar=c(5, 6, 4, 2) + 0.1)
  plot(fROC$cMid,fROC$cHstar-fROC$cLstar,xlab="Location of Gray Zone",ylab="Size of Gray Zone",type="l",lwd=2,cex.axis=1.8,cex.lab=1.8)
  polygon(c(fROC$cMid,rev(fROC$cMid)),c(fROC$cHstar-fROC$cLstar,rep(0,length(fROC$cLstar))),col=gray(0.8))
dev.off()
pdf(file="../PRLettersV2Latex/figures/grayZoneFreq.pdf")
par(mar=c(5, 6, 4, 2) + 0.1)
  plot(fROC$cMid,fROC$grayBest,xlab="Location of Gray Zone",ylab="Proportion of Points in Gray Zone",type="l",lwd=2,cex.axis=2,cex.lab=2)
  polygon(c(fROC$cMid,rev(fROC$cMid)),c(fROC$grayBest,rep(0,length(fROC$cMid))),col=gray(0.8))
#  lines(fROC$cLBest,fROC$grayBest)
#  lines(fROC$cHBest,fROC$grayBest)
  abline(.2,0)
dev.off()
```

Switch to other genes. Make the four files used in figure 5.

```{r}
genes4plot = c("EDF1","CLIP2","ZNF487","COG1")
for (ggName in genes4plot) {
  XX = exp ( exprs(GSE32063_eset)[ggName,] )
  YY = 1 * ( pData(GSE32063_eset)[,"debulking"] == "optimal" )
#  plot(XX,YY,ylab="CLASS LABEL",ylim=c(-.1,1.1),cex=2)
  ROC = grayROC(XX,YY,0)
  fROC = grayROC(XX,YY,.2)
  pdf(file=paste("../PRLettersV2Latex/figures/grayROC_",ggName,".pdf",sep=""))
  grayROCplot(fROC,mainTitle=ggName,set=TRUE,segm=TRUE)
  lines(1-ROC$F0star,1-ROC$F1star)
  dev.off()
}
```

